---
title: "All Tables"
---

```{r, include=FALSE}
library(dplyr)
library(readr)
library(DT)
library(magrittr)

#re-run every now and then to update df - note: might take a few minutes (especially because of IAS)
#source("generate_table_list.R")

table_data <- readr::read_csv("tables/table_list.csv") %>% 
    select(-1) %>% 
    rename(ROWS = RECORDS)

table_summary = table_data %>% 
    group_by(DATABASE) %>% 
    summarize(`Number of tables` = n()) %>% 
    rename(Database = DATABASE) %>% 
    arrange(`Number of tables`)


readLines("tables/table_list.csv") %>% 
  paste0(collapse="\n") %>% 
  openssl::base64_encode() -> encoded
```

The table below includes a list of core database tables available in Vision8 (`V8_NEWYORK_NY_LIVE`), Financial Data Warehouse (`FDW`), and IAS (`IAS`).

Note: for brevity, this table excludes tables with the following schemas: `"dbo", "COMMON", "internal", "DATA_BACKUP", "DataBackup", "CONV", "NYC_STAGE"`. For a comprehensive list of tables that includes these schemas, explore the source data [here](https://github.com/nycdepartmentoffinance/nycdepartmentoffinance.github.io/tree/main/tables).

```{r}
#| echo: false
#| warning: false

knitr::kable(table_summary)
             
```

Use the search bar below to look for a specific type of table (e.g. `parcel_extension`) across the different databases or filter by database explore which tables exist in each resource.

```{r}
#| echo: false
#| warning: false

DT::datatable(table_data, escape = FALSE, filter = 'top', rownames = FALSE,
              options = list(paging = FALSE, scrollCollapse = TRUE,
                             columnDefs = list(
                                 # width limit for table name
                                 list(
                                  targets = 2,
                                  render = JS(
                                    "function(data, type, row, meta) {",
                                    "return type === 'display' && data.length > 15 ?",
                                    "'<span title=\"' + data + '\">' + data.substr(0, 15) + '...</span>' : data;",
                                    "}"))
                             ))
              ) %>% 
    formatStyle(columns = "PREFIX", `max-width` = '15px') %>% 
    formatStyle(columns = "COLS", `max-width` = '15px') %>% 
    formatStyle(columns = "PRIMARY_KEYS", `max-width` = '35px')
```
