---
title: "FDW to Vision Table Crosswalk"

repo-url: https://github.com/nycdepartmentoffinance/nycdepartmentoffinance.github.io
---


```{r, include=FALSE}
library(dplyr)
library(readr)
library(DT)
library(knitr)

#re-run every now and then to update df - note: might take a few minutes (especially because of IAS)
source("generate_fdw_crosswalk.R")

table_data <- readr::read_csv("tables/fdw_crosswalk.csv") %>% 
    select(-1)

table_summary = table_data %>% 
    group_by(DB) %>% 
    summarize(`Number of tables` = n()) %>% 
    rename(Database = DB) %>% 
    arrange(desc(`Number of tables`))

table_data_display = table_data %>% 
    select(-DB) %>% 
    arrange(FDW_NAME)
```

There are currently `r nrow(table_data)` tables within the Financial Data Warehouse (FDW) database from PTS, PDM, and CAMA.

```{r}
#| echo: false
#| warning: false

knitr::kable(table_summary, format = "html", table.attr = "style='width:30%;'")
```


#### Crosswalk

The table below includes all FDW tables, but includes most complete information for those also in the `V8_NEWYORK_NY_LIVE` `REAL_PROP` schema. The primary keys column indicates the identifiers needed to obtain unique records in a database table.


```{r}
#| echo: false
#| warning: false

DT::datatable(table_data_display, escape = FALSE, filter = 'top', rownames = FALSE,
              options = list(paging = FALSE, scrollCollapse = TRUE, bAutoWidth = FALSE,
                                 columnDefs = list(
                                 # width limit for table name
                                 list(
                                  targets = 0,
                                  render = JS(
                                    "function(data, type, row, meta) {",
                                    "return type === 'display' && data.length > 23 ?",
                                    "'<span title=\"' + data + '\">' + data.substr(0, 23) + '...</span>' : data;",
                                    "}"))
                                )),
              
              )
```
