---
title: "PDM/PTS Column Lookup"

repo-url: https://github.com/nycdepartmentoffinance/nycdepartmentoffinance.github.io
---

```{r, include=FALSE}
library(dplyr)
library(readr)
library(DT)
library(magrittr)

source("generate_pts_columns.R")

table_data <- readr::read_csv("tables/pts_columns.csv") %>% 
    select(-1)

```

The table below includes a list of columns from core tables available in the `PDM` tables (including `CONDBASE` and `NEWBASE`) and `PTS` tables (including `ASMT`, `PARDAT`, and `APRVAL`) within the `FDW` database (`FDW3NF`). 

Use the search bar below to look for columns within a particular table (e.g. `ASMT`) or by column name (e.g. `CNS_PID`) to see which tables it appears in.

For comprehensive information on these tables, visit the [PTS Documentation and User Guides](https://nyco365.sharepoint.com/sites/DOF-PropertyDivision/PTS/Shared%20Documents/Forms/AllItems.aspx?id=%2Fsites%2FDOF%2DPropertyDivision%2FPTS%2FShared%20Documents%2FPTS%20Documentation%20and%20User%20Guides&viewid=1f671f8e%2D7586%2D4e62%2Db646%2De0fde9f3da7b&ovuser=32f56fc7%2D5f81%2D4e22%2Da95b%2D15da66513bef%2Cboydclaire%40finance%2Enyc%2Egov&OR=Teams%2DHL&CT=1737575630414&clickparams=eyJBcHBOYW1lIjoiVGVhbXMtRGVza3RvcCIsIkFwcFZlcnNpb24iOiI0OS8yNDEyMDEwMDIxNyIsIkhhc0ZlZGVyYXRlZFVzZXIiOmZhbHNlfQ%3D%3D) internal SharePoint site.

```{r}
#| echo: false
#| warning: false

DT::datatable(table_data, escape = FALSE, filter = 'top', rownames = FALSE,
              options = list(paging = FALSE, scrollCollapse = TRUE,
                             columnDefs = list(
                                 # width limit for table name
                                 list(
                                  targets = 1,
                                  render = JS(
                                    "function(data, type, row, meta) {",
                                    "return type === 'display' && data.length > 15 ?",
                                    "'<span title=\"' + data + '\">' + data.substr(0, 15) + '...</span>' : data;",
                                    "}"))
                                )
                             )
              )
```
