---
title: "Database Table Schema"
---

The table below includes a list of database tables currently available in Vision8 (`V8_NEWYORK_NY_LIVE`), Financial Data Warehouse (`FDW`), and IAS (`IAS`).

Use the search bar below to look for a specific type of table (e.g. `parcel_extension`) across the different databases or filter by database explore which tables exist in each resource. 

To view more column names than display below, hover over the `COLUMN NAMES` value of table you are interested in. 

```{r, include=FALSE}
library(dplyr)
library(readr)
library(DT)

#re-run every now and then to update df - note: might take a few minutes (especially because of IAS)
#source("generate_table_list.R")

table_data <- readr::read_csv("tables/table_list.csv") %>% 
    select(-1)
```

```{r}
#| echo: false
#| warning: false

DT::datatable(table_data, escape = FALSE, filter = 'top', rownames = FALSE,
              options = list(paging = FALSE, scrollCollapse = TRUE,
                             columnDefs = list(
                                 # width limit for column names
                                 list(
                                  targets = 5,
                                  render = JS(
                                    "function(data, type, row, meta) {",
                                    "return type === 'display' && data.length > 15 ?",
                                    "'<span title=\"' + data + '\">' + data.substr(0, 15) + '...</span>' : data;",
                                    "}")),
                                 # width limit for table name
                                 list(
                                  targets = 2,
                                  render = JS(
                                    "function(data, type, row, meta) {",
                                    "return type === 'display' && data.length > 20 ?",
                                    "'<span title=\"' + data + '\">' + data.substr(0, 20) + '...</span>' : data;",
                                    "}"))
                                ))
              ) %>% 
    formatStyle(columns = "PREFIX", `max-width` = '60px') %>% 
    formatStyle(columns = "COLS", `max-width` = '60px')
```
