---
title: "Co-op Sales Model"

repo-url: https://github.com/nycdepartmentoffinance/nycdepartmentoffinance.github.io
---

```{r, include=FALSE}
# read env file to get proxy set up to work
readRenviron("C:/Users/BoydClaire/.Renviron")
options(scipen=99999)

library(readxl)
library(stringr)
library(readr)
library(ggplot2)
library(dplyr)
library(sf)

library(maps)
library(leaflet)
library(scales)
library(DT)
```

```{r functions, include=FALSE, cache=TRUE}
# color palettes to use across plots/maps
color_pal_r <- function(x) {
  sapply(x, function(val) {
    if (is.na(val)) return("#999999")  # gray for NA

    if (val >= 0.95) {
      return("#00FF00")  # pure green, consistent with others
    } else if (val >= 0.7 && val < 0.95) {
      # Green to Yellow: R goes from 0 → 1, G stays at 1
      intensity <- rescale(val, from = c(0.7, 1), to = c(1, 0))  # reverse scale
      return(rgb(intensity, 1, 0))  # green to yellow
    } else if (val >= 0.4 && val < 0.7) {
      # Yellow to Red: G goes from 1 → 0, R stays at 1
      intensity <- rescale(val, from = c(0.4, 0.7), to = c(0, 1))
      return(rgb(1, intensity, 0))  # yellow to red
    } else {
      return("#FF0000")  # solid red for <0.4
    }
  })
}

color_pal_cod <- function(x) {
  sapply(x, function(val) {
    if (is.na(val)) return("#999999")  # gray for NA

    if (val >= 5 & val <= 10) {
      return("#00FF00")  # Still using green for the "ideal" zone
    } else if (val < 5) {
      # Values < 5: interpolate from yellow (at 5) to orange (at 0)
      # Yellow = rgb(1, 1, 0), Orange = rgb(1, 0.5, 0)
      intensity <- scales::rescale(val, from = c(0, 5), to = c(0.5, 1))  # Green component goes from 0.5 to 1
      return(rgb(1, intensity, 0))
    } else {
      # Interpolate from yellow (10) to red (20), capped at red for >20
      capped_val <- min(val, 20)
      intensity <- rescale(capped_val, from = c(10, 20), to = c(1, 0))  # G goes from 1 to 0
      return(rgb(1, intensity, 0))  # Yellow to Red
    }
  })
}

color_pal_prd <- function(x) {
  sapply(x, function(val) {
    if (is.na(val)) return("#999999")  # gray for NA

    if (val >= 0.98 & val <= 1.03) {
      return("#00FF00")  # ideal green
    } else if (val < 0.98) {
      # Light blue (0.98) to dark blue (~0.90): blue stays 1, R & G go from 0.8 to 0
      capped_val <- max(val, 0.90)
      intensity <- scales::rescale(capped_val, from = c(0.90, 0.98), to = c(0, 0.8))
      return(rgb(intensity, intensity, 1))  # light to dark blue
    } else {
      # Light red (1.03) to dark red (~1.10): red stays 1, G & B go from 0.8 to 0
      capped_val <- min(val, 1.10)
      intensity <- scales::rescale(capped_val, from = c(1.03, 1.10), to = c(0.8, 0))
      return(rgb(1, intensity, intensity))  # light to dark red
    }
  })
}

color_pal_sales_ratio <- function(x) {
  sapply(x, function(val) {
    if (is.na(val)) return("#999999")  # gray for NA

    if (val >= 0.9 & val <= 1.1) {
      return("green")  # Ideal range
    } else if (val < 0.9) {
      # Light blue (0.9) to dark blue (0.6): blue stays 1, R & G fade from 0.6 to 0
      capped_val <- max(val, 0.6)
      intensity <- scales::rescale(capped_val, from = c(0.6, 0.9), to = c(0, 0.6))
      return(rgb(intensity, intensity, 1))  # Light blue to dark blue
    } else {
      # Light red (1.1) to dark red (1.4): red stays 1, G & B fade from 0.6 to 0
      capped_val <- min(val, 1.4)
      intensity <- scales::rescale(capped_val, from = c(1.1, 1.4), to = c(0.6, 0))
      return(rgb(1, intensity, intensity))  # Light red to dark red
    }
  })
}


color_pal_grosscon <- function(x){
 
  clamped <- pmin(pmax(x, 200000), 5000000)
    
  pal = colorNumeric(
      palette = "Blues",  # You can also use c("#deebf7", "#3182bd") for more control
      domain = c(200000, 5000000),  # fixed scale
      na.color = "#999999"  # gray for NA
  )
  return(pal(clamped))
}

#table formatting
format_metrics_table <- function(data, palettes, opacity = 0.2, exclude_cols = c("ci", "gini")) {
  # Add color columns using provided palettes and opacity
  data <- data %>%
    mutate(across(
      .cols = names(palettes),
      .fns = ~ hex_with_opacity(palettes[[cur_column()]](.x), opacity),
      .names = "color_{.col}"
    ))
  
  # Determine which columns to show
  visible_cols <- setdiff(names(data), c(paste0("color_", names(palettes)), exclude_cols))
  
  # Build the datatable
  dt <- datatable(
    data[, visible_cols],
    options = list(pageLength = 6, scrollX = TRUE)
  )
  
  # Apply conditional formatting
  for (metric in names(palettes)) {
    dt <- dt %>% formatStyle(
      metric,
      backgroundColor = styleEqual(
        data[[metric]],
        data[[paste0("color_", metric)]]
      )
    )
  }
  
  return(dt)
}


# map functions
create_metrics_map <- function(data, by, center_lng = -73.96, center_lat = 40.78, zoom_level = 12) {
  
  leaflet(data = data, height=800) %>%
    addTiles() %>%
    setView(center_lng, center_lat, zoom = zoom_level) %>%
    addProviderTiles("CartoDB.Positron") %>% 
    {
      map <- .
      
          # Create popup content using get() to dynamically access the ID field
          popup_content <- eval(parse(text = paste0(
            "~paste('<strong>", by, ":</strong> ', get('", by, "'), '<br>',",
            "'<strong>n:</strong> ', n, '<br>',",
            "'<strong>median_sales_ratio:</strong> ', median_sales_ratio, '<br>',",
            "'<strong>r:</strong> ', r, '<br>',",
            "'<strong>spearmans:</strong> ', spearmans, '<br>',",
            "'<strong>cod:</strong> ', cod, '<br>',",
            "'<strong>prd:</strong> ', prd, '<br>',",
            "'<strong>prb:</strong> ', prb, '<br>',",
            "'<strong>clapp:</strong> ', clapp, '<br>',",
            "'<strong>ki:</strong> ', ki, '<br>',",
            "'<strong>mki:</strong> ', mki, '<br>')"
          )))
    
      
      # Define function to add polygon layers
      add_polygon_layer <- function(m, group_name, var_name, color_fn) {
        m %>% addPolygons(
          group = group_name,
          fillColor = ~color_fn(get(var_name)),
          fillOpacity = 0.7,
          color = "white",
          weight = 1,
          popup = popup_content
        )
      }
      
      # Add all layers using the function
      map %>%
        add_polygon_layer("By R", "r", color_pal_r) %>%
        add_polygon_layer("By Spearman's Rank", "spearmans", color_pal_r) %>%
        add_polygon_layer("By COD", "cod", color_pal_cod) %>%
        add_polygon_layer("By PRD", "prd", color_pal_prd) %>%
        add_polygon_layer("By MKI", "mki", color_pal_r)
    } %>%
    
    # Add control to switch between layers
    addLayersControl(
      baseGroups = c("By R", "By Spearman's Rank","By COD", "By PRD", "By MKI"),
      options = layersControlOptions(collapsed = FALSE)
    ) %>%
    
    # Only show one group by default
    hideGroup(c("By Spearman's Rank","By COD", "By PRD","By MKI"))
}

create_point_map <- function(data, 
                             dependent_variables,
                             layers =
                              list(
                                list(group = "By Sale Price", var = "GROSSCON", color_pal = color_pal_grosscon),
                                list(group = "By Predicted Sale Price", var = "prediction", color_pal = color_pal_grosscon),
                                list(group = "By Sales Ratio", var = "sales_ratio", color_pal = color_pal_sales_ratio)
                              ), 
                             center_lng = -73.96, 
                             center_lat = 40.78, 
                             zoom_level = 13) {

  # Required top-level fields, formatted
  popup_content <- sapply(seq_len(nrow(data)), function(i) {
    row <- data[i, ]

    main_info <- paste0(
      "<strong>PID:</strong> ", row[["PID"]], "<br>",
      "<strong>SALEKEY:</strong> ", row[["SALEKEY"]], "<br>",
      "<strong>Sale Price (real value):</strong> $", scales::comma(round(row[["GROSSCON"]])), "<br>",
      "<strong>Predicted Sale Price:</strong> $", scales::comma(round(row[["prediction"]])), "<br>",
      "<strong>Sales Ratio:</strong> ", round(row[["sales_ratio"]], 3), "<br><br>"
    )

    dependent_vars <- paste0(
      "<strong>Dependent Variables:</strong><br>",
      paste0(
        lapply(dependent_variables, function(var) {
          paste0("<strong>", var, ":</strong> ", row[[var]])
        }),
        collapse = "<br>"
      )
    )

    paste0(main_info, dependent_vars)
  })

  map <- leaflet(data = data, height = 800) %>%
    addTiles() %>%
    setView(lng = center_lng, lat = center_lat, zoom = zoom_level) %>%
    addProviderTiles("CartoDB.Positron")

  for (layer in layers) {
    group_name <- layer$group
    var_name <- layer$var
    color_fn <- layer$color_pal

    map <- map %>%
      addCircleMarkers(
        data = data,
        group = group_name,
        fillColor = ~color_fn(get(var_name)),
        color = "black",
        weight = 1,
        radius = 6,
        fillOpacity = 0.7,
        stroke = TRUE,
        popup = popup_content
      )
  }

  map %>%
    addLayersControl(
      baseGroups = sapply(layers, function(l) l$group),
      options = layersControlOptions(collapsed = FALSE)
    ) %>%
    hideGroup(sapply(layers[-1], function(l) l$group))
}


hex_with_opacity <- function(hex, alpha = 1) {
  hex <- gsub("#", "", hex)  # Remove '#' if present

  # Convert alpha (0–1) to 2-digit hex
  alpha_hex <- toupper(sprintf("%02X", round(alpha * 255)))

  paste0("#", hex, alpha_hex)
}

```

```{r data_ingestion, include=FALSE, cache=TRUE}
source_folder = "C:/Users/BoydClaire/workspace_local/coop_sales_model/model_results/grosscon_psqft_boro_2pt/mixed_model_block"

# model metrics
boro_metrics = readr::read_csv(glue::glue("{source_folder}/metrics.csv")) %>% 
    mutate(across(
        .cols = -c(boro, predicted, actual, n, accuracy, uniformity, vertical_equity), 
        .fns = ~round(.x, 4)
      )) %>% 
    select(-predicted, -actual)

neighborhood_metrics = readr::read_csv(glue::glue("{source_folder}/neighborhood_metrics.csv")) 

block_metrics = readr::read_csv(glue::glue("{source_folder}/manhattan_block_metrics.csv")) %>% 
    mutate(block = as.numeric(block))

train_df = readr::read_csv(glue::glue("{source_folder}/train_with_yhat.csv"))

# shapefiles // geographic transformations
neighborhood_shapefile = sf::st_read("data/DOF_NBHD_Dissolved/DOF_NBHD_Dissolved.shp") %>% 
    st_transform(crs=4326) %>% 
    left_join(neighborhood_metrics, by=c("CODE"="NEIGHBORHOOD"))

block_shapefile = sf::st_read("data/TAX_BLOCK_POLYGON_LAYER/TAX_BLOCK_POLYGON_LAYER.shp") %>% 
    st_transform(crs=4326) %>% 
    filter(BORO == "1") %>% 
    select(BLOCK, geometry) %>% 
    left_join(block_metrics, by=c("BLOCK"="block"))

train <- sf::st_as_sf(train_df, coords= c("xcoord","ycoord"), crs=2263) %>% 
    st_transform(crs=4326) 
```

Using sales to predict the sale price of co-ops across NYC, the most successful model to date is the following:

```{r, results='hide'}
log_grosscon ~ NEIGHBORHOOD + xcoord_norm + ycoord_norm + BC + 
    park_dist + cultural_facility_dist + 
    (1 | boro) + 
    (1 + stories + affordable | block) + 
    (1 + BEDROOMS + BATHROOMS + floor + log_sqft | COOPID)
```

### City-wide

To assess variation in model performance across the city, key assessment accuracy metrics are summarized at the borough level in the table below. For an overview of how all the visuals are shaded, see [notes](#notes).

```{r, echo=FALSE}
metric_palettes <- list(
  r = color_pal_r,
  spearmans = color_pal_r,
  cod = color_pal_cod,
  prd = color_pal_prd,
  mki = color_pal_r
)

format_metrics_table(boro_metrics, palettes = metric_palettes, opacity = 0.2)
```

<br>

The interactive map below visualizes key assessment accuracy metrics at the neighborhood level across New York City. Toggle between the different layers which shade neighborhoods by different performance indicators, such as: R, Spearman's Rank, Coefficient of Dispersion (COD), Price-related Differential (PRD), and Modified Kakwani Index (MKI). Clicking on a neighborhood reveals a popup with detailed statistics for that area.

***Note:** Neighborhoods with less than 10 sales used to train the model do not display any accuracy metrics.*

```{r, cache=TRUE}
create_metrics_map(
  data = neighborhood_shapefile, 
  by = "NBHD_NAME",
  center_lng = -73.99, 
  center_lat = 40.76, 
  zoom_level = 11
)
```

### Manhattan

The map below displays the model's performance in Manhattan at the block level. Toggle between the layers to display different accuracy metrics or click on a block to more carefully inspect how well the model performs on sales on that block.

```{r, cache=TRUE}
create_metrics_map(
  data = block_shapefile, 
  by = "BLOCK",
  center_lng = -73.96, 
  center_lat = 40.78, 
  zoom_level = 12
)
```

<br> This map shows a breakdown of the `r scales::comma(nrow(train[train$boro == 1,]))` Manhattan sales used to train the model, shaded by either the sale price, predicted sale price, or the sales ratio. Clicking on a sale reveals a popup with the sale's characteristics used by the model to determined it's predicted value (dependent variables).

```{r, cache=TRUE}
create_point_map(
  data = train[train$boro == 1,],
  dependent_variables = c("BC", "park_dist", "cultural_facility_dist", 
    "stories", "affordable", "BEDROOMS", "BATHROOMS", "floor", "log_sqft"),
  center_lng = -73.96, 
  center_lat = 40.78, 
  zoom_level = 12
)
```

### Notes

The table below summarizes how the visuals are all uniformly shaded.

```{r, echo=FALSE}
# Create the summary data frame
color_pal_summary <- data.frame(
  palette = c(
    "color_pal_r", 
    "color_pal_cod", 
    "color_pal_prd", 
    "color_pal_sales_ratio", 
    "color_pal_grosscon"
  ),
  green_range = c(
    "≥ 0.95", 
    "5–10", 
    "0.98–1.03", 
    "0.9–1.1", 
    "N/A"
  ),
  below_range = c(
    "0.4–0.95: Red to Yellow to Green", 
    "0–5: Red to Yellow", 
    "0.90–0.98: Dark Blue to Light Blue (Progressive)", 
    "0.6–0.9: Dark Blue to Light Blue (Progressive)", 
    "N/A"
  ),
  above_range = c(
    "N/A", 
    "10–20: Yellow to Red", 
    "1.03–1.10: Light to Dark Red (Regressive)", 
    "1.1–1.4: Light to Dark Red (Regressive)", 
    "Blue gradient ($200K–$5M)"
  ),
  stringsAsFactors = FALSE
)

# Display table with minimal formatting
datatable(
  color_pal_summary,
  options = list(
    dom = 't',
    ordering = FALSE
  ),
  rownames = FALSE
)

```



