---
title: "Co-op Sales Model"

repo-url: https://github.com/nycdepartmentoffinance/nycdepartmentoffinance.github.io
---

```{r, include=FALSE}
# read env file to get proxy set up to work
readRenviron("Data/.Renviron")
options(scipen=99999)

library(readxl)
library(stringr)
library(readr)
library(ggplot2)
library(dplyr)
library(sf)
library(rmarkdown)
library(maps)
library(leaflet)
library(scales)
library(DT)
```

```{r functions, include=FALSE, cache=TRUE}
# color palettes to use across plots/maps
color_pal_r <- function(x) {
  sapply(x, function(val) {
    if (is.na(val)) return("#999999")  # gray for NA

    if (val >= 0.95) {
      return("#00FF00")  # pure green, consistent with others
    } else if (val >= 0.7 && val < 0.95) {
      # Green to Yellow: R goes from 0 → 1, G stays at 1
      intensity <- rescale(val, from = c(0.7, 1), to = c(1, 0))  # reverse scale
      return(rgb(intensity, 1, 0))  # green to yellow
    } else if (val >= 0.4 && val < 0.7) {
      # Yellow to Red: G goes from 1 → 0, R stays at 1
      intensity <- rescale(val, from = c(0.4, 0.7), to = c(0, 1))
      return(rgb(1, intensity, 0))  # yellow to red
    } else {
      return("#FF0000")  # solid red for <0.4
    }
  })
}

color_pal_cod <- function(x) {
  sapply(x, function(val) {
    if (is.na(val)) return("#999999")  # gray for NA

    if (val >= 5 & val <= 15) {
      return("#00FF00")  # Still using green for the "ideal" zone
    } else if (val < 5) {
      # Values < 5: interpolate from yellow (at 5) to orange (at 0)
      # Yellow = rgb(1, 1, 0), Orange = rgb(1, 0.5, 0)
      intensity <- scales::rescale(val, from = c(0, 5), to = c(0.5, 1))  # Green component goes from 0.5 to 1
      return(rgb(1, intensity, 0))
    } else {
      # Interpolate from yellow (10) to red (20), capped at red for >20
      capped_val <- min(val, 20)
      intensity <- rescale(capped_val, from = c(15, 20), to = c(1, 0))  # G goes from 1 to 0
      return(rgb(1, intensity, 0))  # Yellow to Red
    }
  })
}

color_pal_prd <- function(x) {
  sapply(x, function(val) {
    if (is.na(val)) return("#999999")  # gray for NA

    if (val >= 0.98 & val <= 1.03) {
      return("#00FF00")  # ideal green
    } else if (val < 0.98) {
      # Light blue (0.98) to dark blue (~0.90): blue stays 1, R & G go from 0.8 to 0
      capped_val <- max(val, 0.90)
      intensity <- scales::rescale(capped_val, from = c(0.90, 0.98), to = c(0, 0.8))
      return(rgb(intensity, intensity, 1))  # light to dark blue
    } else {
      # Light red (1.03) to dark red (~1.10): red stays 1, G & B go from 0.8 to 0
      capped_val <- min(val, 1.10)
      intensity <- scales::rescale(capped_val, from = c(1.03, 1.10), to = c(0.8, 0))
      return(rgb(1, intensity, intensity))  # light to dark red
    }
  })
}

color_pal_sales_ratio <- function(x) {
  sapply(x, function(val) {
    if (is.na(val)) return("#999999")  # gray for NA

    if (val >= 0.9 & val <= 1.1) {
      return("green")  # Ideal range
    } else if (val < 0.9) {
      # Light blue (0.9) to dark blue (0.6): blue stays 1, R & G fade from 0.6 to 0
      capped_val <- max(val, 0.6)
      intensity <- scales::rescale(capped_val, from = c(0.6, 0.9), to = c(0, 0.6))
      return(rgb(intensity, intensity, 1))  # Light blue to dark blue
    } else {
      # Light red (1.1) to dark red (1.4): red stays 1, G & B fade from 0.6 to 0
      capped_val <- min(val, 1.4)
      intensity <- scales::rescale(capped_val, from = c(1.1, 1.4), to = c(0.6, 0))
      return(rgb(1, intensity, intensity))  # Light red to dark red
    }
  })
}


color_pal_grosscon <- function(x){
 
  clamped <- pmin(pmax(x, 200000), 5000000)
    
  pal = colorNumeric(
      palette = "Blues",  # You can also use c("#deebf7", "#3182bd") for more control
      domain = c(200000, 5000000),  # fixed scale
      na.color = "#999999"  # gray for NA
  )
  return(pal(clamped))
}

color_pal_mki <- function(x) {
  sapply(x, function(val) {
    if (is.na(val)) return("#999999")  # gray for NA

    if (val > 1.4) {
      return("#FF0000")  # solid red for extreme high values
    } else if (val > 1.1 && val <= 1.4) {
      # Yellow to Red for values >1.1 to 1.4
      intensity <- rescale(val, from = c(1.1, 1.4), to = c(1, 0))
      return(rgb(1, intensity, 0))  # yellow to red
    } else if (val >= 0.9 && val <= 1.1) {
      return("#00FF00")  # solid yellow as center plateau
    } else if (val >= 0.6 && val < 0.9) {
      # Yellow to Red from 0.9 to 0.6
      intensity <- rescale(val, from = c(0.9, 0.6), to = c(1, 0))
      return(rgb(1, intensity, 0))  # yellow to red
    } else {
      return("#FF0000")  # solid red for <0.6
    }
  })
}


#table formatting
format_metrics_table <- function(data, palettes, opacity = 0.2, exclude_cols = c("ci", "gini")) {
  # Add color columns using provided palettes and opacity
  data <- data %>%
    mutate(across(
      .cols = names(palettes),
      .fns = ~ hex_with_opacity(palettes[[cur_column()]](.x), opacity),
      .names = "color_{.col}"
    ))
  
  # Determine which columns to show
  visible_cols <- setdiff(names(data), c(paste0("color_", names(palettes)), exclude_cols))
  
  # Build the datatable
  dt <- datatable(
    data[, visible_cols],
    options = list(pageLength = 6, scrollX = TRUE)
  )
  
  # Apply conditional formatting
  for (metric in names(palettes)) {
    dt <- dt %>% formatStyle(
      metric,
      backgroundColor = styleEqual(
        data[[metric]],
        data[[paste0("color_", metric)]]
      )
    )
  }
  
  return(dt)
}


# map functions
create_metrics_map <- function(data, by, center_lng = -73.96, center_lat = 40.78, zoom_level = 12) {
  
  leaflet(data = data, height=800) %>%
    addTiles() %>%
    setView(center_lng, center_lat, zoom = zoom_level) %>%
    addProviderTiles("CartoDB.Positron") %>% 
    {
      map <- .
      
          # Create popup content using get() to dynamically access the ID field
          popup_content <- eval(parse(text = paste0(
            "~paste('<strong>", by, ":</strong> ', get('", by, "'), '<br>',",
            "'<strong>n:</strong> ', n, '<br>',",
            "'<strong>median_sales_ratio:</strong> ', median_sales_ratio, '<br>',",
            "'<strong>r:</strong> ', r, '<br>',",
            "'<strong>spearmans:</strong> ', spearmans, '<br>',",
            "'<strong>cod:</strong> ', cod, '<br>',",
            "'<strong>prd:</strong> ', prd, '<br>',",
            "'<strong>prb:</strong> ', prb, '<br>',",
            "'<strong>clapp:</strong> ', clapp, '<br>',",
            "'<strong>ki:</strong> ', ki, '<br>',",
            "'<strong>mki:</strong> ', mki, '<br>')"
          )))
    
      
      # Define function to add polygon layers
      add_polygon_layer <- function(m, group_name, var_name, color_fn) {
        m %>% addPolygons(
          group = group_name,
          fillColor = ~color_fn(get(var_name)),
          fillOpacity = 0.7,
          color = "white",
          weight = 1,
          popup = popup_content
        )
      }
      
      # Add all layers using the function
      map %>%
        add_polygon_layer("By R", "r", color_pal_r) %>%
        add_polygon_layer("By Spearman's Rank", "spearmans", color_pal_r) %>%
        add_polygon_layer("By COD", "cod", color_pal_cod) %>%
        add_polygon_layer("By PRD", "prd", color_pal_prd) %>%
        add_polygon_layer("By MKI", "mki", color_pal_mki)
    } %>%
    
    # Add control to switch between layers
    addLayersControl(
      baseGroups = c("By R", "By Spearman's Rank","By COD", "By PRD", "By MKI"),
      options = layersControlOptions(collapsed = FALSE)
    ) %>%
    
    # Only show one group by default
    hideGroup(c("By Spearman's Rank","By COD", "By PRD","By MKI"))
}

create_point_map <- function(data, 
                             dependent_variables,
                             layers =
                              list(
                                list(group = "By Sales Ratio", var = "sales_ratio", color_pal = color_pal_sales_ratio),
                                list(group = "By Sale Price", var = "GROSSCON", color_pal = color_pal_grosscon),
                                list(group = "By Predicted Sale Price", var = "prediction", color_pal = color_pal_grosscon)
                              ), 
                             center_lng = -73.96, 
                             center_lat = 40.78, 
                             zoom_level = 13) {

  # Required top-level fields, formatted
  popup_content <- sapply(seq_len(nrow(data)), function(i) {
    row <- data[i, ]

    main_info <- paste0(
      "<strong>PID:</strong> ", row[["PID"]], "<br>",
      "<strong>COOPUNITKEY:</strong> ", row[["COOPUNITKEY"]], "<br>",
      "<strong>ADDRESS:</strong> ", row[["ADRNO"]], " ", row[["ADRSTR"]], "<br>",
      "<strong>SALEKEY:</strong> ", row[["SALEKEY"]], "<br>",
      "<strong>Sale Price (real value):</strong> $", scales::comma(round(row[["GROSSCON"]])), "<br>",
      "<strong>Predicted Sale Price:</strong> $", scales::comma(round(row[["prediction"]])), "<br>",
      "<strong>Sales Ratio:</strong> ", round(row[["sales_ratio"]], 3), "<br><br>"
    )

    dependent_vars <- paste0(
      "<strong>Dependent Variables:</strong><br>",
      paste0(
        lapply(dependent_variables, function(var) {
          paste0("<strong>", var, ":</strong> ", row[[var]])
        }),
        collapse = "<br>"
      )
    )

    paste0(main_info, dependent_vars)
  })

  map <- leaflet(data = data, height = 800) %>%
    addTiles() %>%
    setView(lng = center_lng, lat = center_lat, zoom = zoom_level) %>%
    addProviderTiles("CartoDB.Positron")

  for (layer in layers) {
    group_name <- layer$group
    var_name <- layer$var
    color_fn <- layer$color_pal

    map <- map %>%
      addCircleMarkers(
        data = data,
        group = group_name,
        fillColor = ~color_fn(get(var_name)),
        color = "black",
        weight = 1,
        radius = 6,
        fillOpacity = 0.7,
        stroke = TRUE,
        popup = popup_content
      )
  }

  map %>%
    addLayersControl(
      baseGroups = sapply(layers, function(l) l$group),
      options = layersControlOptions(collapsed = FALSE)
    ) %>%
    hideGroup(sapply(layers[-1], function(l) l$group))
}


create_building_map <- function(data, 
                             display_variables,
                             layers, 
                             center_lng = -73.96, 
                             center_lat = 40.78, 
                             zoom_level = 13) {

  # Required top-level fields, formatted
  popup_content <- sapply(seq_len(nrow(data)), function(i) {
    row <- data[i, ]

    display_vars <- paste0(
      paste0(
        lapply(display_variables, function(var) {
          paste0("<strong>", var, ":</strong> ", row[[var]])
        }),
        collapse = "<br>"
      )
    )

    display_vars
  })

  map <- leaflet(data = data, height = 800) %>%
    addTiles() %>%
    setView(lng = center_lng, lat = center_lat, zoom = zoom_level) %>%
    addProviderTiles("CartoDB.Positron")

  for (layer in layers) {
    group_name <- layer$group
    var_name <- layer$var
    color_fn <- layer$color_pal

    map <- map %>%
      addCircleMarkers(
        data = data,
        group = group_name,
        fillColor = ~color_fn(get(var_name)),
        color = "black",
        weight = 1,
        radius = 6,
        fillOpacity = 0.7,
        stroke = TRUE,
        popup = popup_content
      )
  }

  map %>%
    addLayersControl(
      baseGroups = sapply(layers, function(l) l$group),
      options = layersControlOptions(collapsed = FALSE)
    ) %>%
    hideGroup(sapply(layers[-1], function(l) l$group))
}



hex_with_opacity <- function(hex, alpha = 1) {
  hex <- gsub("#", "", hex)  # Remove '#' if present

  # Convert alpha (0–1) to 2-digit hex
  alpha_hex <- toupper(sprintf("%02X", round(alpha * 255)))

  paste0("#", hex, alpha_hex)
}

```

```{r data_ingestion, include=FALSE}
source_folder = "Data/Coop/MixedModel"

# model metrics
boro_metrics = readr::read_csv(glue::glue("{source_folder}/metrics.csv")) %>% 
    mutate(across(
        .cols = -c(boro, boro_name, predicted, actual, n, accuracy, uniformity, vertical_equity), 
        .fns = ~round(.x, 4)
      )) %>% 
    select(-predicted, -actual)

# comps_boro_metrics = readr::read_csv(glue::glue("{source_folder}/comps_metrics.csv")) %>% 
#     mutate(across(
#         .cols = -c(boro, predicted, actual, n, accuracy, uniformity, vertical_equity), 
#         .fns = ~round(.x, 4)
#       )) %>% 
#     select(-predicted, -actual)

neighborhood_metrics = readr::read_csv(glue::glue("{source_folder}/neighborhood_metrics.csv")) 

block_metrics = readr::read_csv(glue::glue("{source_folder}/manhattan_block_metrics.csv")) %>% 
    mutate(block = as.numeric(block))

neighborhood_decile_metrics = readr::read_csv(glue::glue("{source_folder}/neighborhood_decile_metrics.csv")) %>% 
    mutate(across(
        .cols = -c(boro, boro_name, NBHD_NAME, NEIGHBORHOOD, cutoff, decile, predicted, actual, n, accuracy, uniformity, vertical_equity), 
        .fns = ~round(.x, 4)
      )) %>% 
    select(-predicted, -actual)

boro_decile_metrics = readr::read_csv(glue::glue("{source_folder}/boro_decile_metrics.csv"))%>% 
    mutate(across(
        .cols = -c(boro, boro_name, decile, predicted, cutoff, actual, n, accuracy, uniformity, vertical_equity), 
        .fns = ~round(.x, 4)
      )) %>% 
    select(-predicted, -actual)

train_df = readr::read_csv(glue::glue("{source_folder}/train_with_yhat.csv"))

# shapefiles // geographic transformations
neighborhood_shapefile = sf::st_read("data/DOF_NBHD_Dissolved/DOF_NBHD_Dissolved.shp") %>% 
    st_transform(crs=4326) %>% 
    left_join(neighborhood_metrics, by=c("CODE"="NEIGHBORHOOD"))

block_shapefile = sf::st_read("data/TAX_BLOCK_POLYGON_LAYER/TAX_BLOCK_POLYGON_LAYER.shp") %>% 
    st_transform(crs=4326) %>% 
    filter(BORO == "1") %>% 
    select(BLOCK, geometry) %>% 
    left_join(block_metrics, by=c("BLOCK"="block"))

train <- sf::st_as_sf(train_df, coords= c("xcoord","ycoord"), crs=2263) %>% 
    st_transform(crs=4326) 
```

## Mixed Model

Using sales from 2022-2024 to predict the sale price of co-ops across NYC, the most successful mixed model (or hierarchical linear model) to date is the following:

```{r, results='hide'}
log_grosscon ~ NEIGHBORHOOD + BC + xcoord + ycoord + affordable +
    park_dist + cultural_facility_dist + subway_dist + gym + doorman + laundry +
    (0 + stories_norm | block) +
    (1 + BEDROOMS_norm + BATHROOMS_norm + log_floor_norm + log_sqft_norm | COOPID)
```

### City-wide

To assess variation in model performance across the city, key assessment accuracy metrics are summarized at the borough level in the table below. For an overview of how all the visuals are shaded, see [notes](#notes).

```{r, echo=FALSE}
metric_palettes <- list(
  r = color_pal_r,
  spearmans = color_pal_r,
  cod = color_pal_cod,
  prd = color_pal_prd,
  mki = color_pal_mki
)

format_metrics_table(boro_metrics, palettes = metric_palettes, opacity = 0.2)
```

<br>

The interactive map below visualizes key assessment accuracy metrics at the neighborhood level across New York City. Toggle between the different layers which shade neighborhoods by different performance indicators, such as: R, Spearman's Rank, Coefficient of Dispersion (COD), Price-related Differential (PRD), and Modified Kakwani Index (MKI). Clicking on a neighborhood reveals a popup with detailed statistics for that area.

***Note:** Neighborhoods with less than 10 sales used to train the model do not display any accuracy metrics.*

```{r}
create_metrics_map(
  data = neighborhood_shapefile, 
  by = "NBHD_NAME",
  center_lng = -73.99, 
  center_lat = 40.76, 
  zoom_level = 11
)
```

### Manhattan

The map below displays the model's performance in Manhattan at the block level. Toggle between the layers to display different accuracy metrics or click on a block to more carefully inspect how well the model performs on sales on that block.

```{r}
create_metrics_map(
  data = block_shapefile, 
  by = "BLOCK",
  center_lng = -73.96, 
  center_lat = 40.78, 
  zoom_level = 12
)
```

<br> This map shows a breakdown of the `r scales::comma(nrow(train[train$boro == 1,]))` Manhattan sales used to train the model, shaded by either the sale price, predicted sale price, or the sales ratio. Clicking on a sale reveals a popup with the sale's characteristics used by the model to determined it's predicted value (dependent variables).

```{r}
create_point_map(
  data = train[train$boro == 1,],
  dependent_variables = c("NBHD_NAME", "BC", "park_dist", "cultural_facility_dist", "subway_dist",
    "affordable","gym", "laundry", "doorman","stories","floor","BEDROOMS","BATHROOMS","log_sqft"),
  center_lng = -73.96, 
  center_lat = 40.78, 
  zoom_level = 12
)
```


### By Sale Price Decile

The two tables below summarize the model's performance by actual sale price (GROSSCON) decile, which are calculated at the boro level (e.g. the upper threshold of bottom 10% of sales in Manhattan is \$400,000 while it is \$175,000 in Staten Island).

#### Borough Metrics by Sale Price Decile

```{r, warning=FALSE, echo=FALSE}
# Ensure decile is a factor with levels in the desired order
# Convert borough names to numeric values
boro_decile_metrics$boro_num = as.numeric(boro_decile_metrics$boro)

# Define borough colors
boro_colors <- c("red", "blue", "maroon", "purple", "darkgreen")
colors <- boro_colors[boro_decile_metrics$boro_num]  # Assuming boro is a number from 1 to 5

# Create the scatter plot without x-axis ticks
# The points will be drawn after the shaded region
plot(boro_decile_metrics$decile, boro_decile_metrics$mki,
     col = colors,
     pch = 19,  # filled circles
     xaxt = "n",  # suppress x-axis
     xlab = "Decile (of GROSSCON)", ylab = "MKI", main = "MKI by Decile and Borough",
     ylim = c(0, 2)
)

# Add custom x-axis with decile labels 1 to 10
axis(1, at = 1:10, labels = 1:10)

# Add shaded region between y = 0.9 and y = 1.1 (polygon behind the points)
polygon(c(0:11, rev(0:11)), c(rep(0.9, 12), rev(rep(1.1, 12))),
        col = "lightgreen", border = NA)  # Shaded region

# Add points (text labels) for each borough
points(boro_decile_metrics$decile, boro_decile_metrics$mki, 
       col = colors, pch = 19)

# Optional: Add a legend for boroughs
legend("top", legend = c("MN", "BX", "BK", "QN", "SI"),
       col = boro_colors, pch = 19, horiz = TRUE)

```

Here is the source data to more easily explore the raw data from the plot above.

```{r, echo=FALSE}
metric_palettes <- list(
  r = color_pal_r,
  spearmans = color_pal_r,
  cod = color_pal_cod,
  prd = color_pal_prd,
  mki = color_pal_mki
)

format_metrics_table(boro_decile_metrics, palettes = metric_palettes, opacity = 0.2)
```

#### Neighborhood Metrics by Sale Price Decile

The variation of MKI by neighborhood is generally consistent across deciles of GROSSCON, largely ranging around 1 (ideal value), though the mean of most deciles is below 1. There are more outliers in the 4-8 decile ranges, indicating that there might be a poorer model fit in the middle ranges of the dataset.

```{r, warning=FALSE, echo=FALSE}
# Ensure decile is a factor with levels in the desired order
neighborhood_decile_metrics$decile <- factor(neighborhood_decile_metrics$decile,
                                              levels = as.character(1:10))

# Plot
boxplot(mki ~ decile, data = neighborhood_decile_metrics,
        xlab = "Decile (of GROSSCON)", ylab = "MKI", main = "Neighborhood MKI by GROSSCON Decile")

# Add horizontal line at y = 1
abline(h = 1, col = "green", lty = 2, lwd = 2)
```

The table below contains the source data used to generate the plot above.

```{r, echo=FALSE}
metric_palettes <- list(
  r = color_pal_r,
  spearmans = color_pal_r,
  cod = color_pal_cod,
  prd = color_pal_prd,
  mki = color_pal_mki
)

neighborhood_decile_metrics_trim = neighborhood_decile_metrics %>% 
    select(-NEIGHBORHOOD, -boro_name)

format_metrics_table(neighborhood_decile_metrics_trim, palettes = metric_palettes, opacity = 0.2)
```


### Sale Price Distribution by COOP building

The point map below displays the standard deviation of the sale prices in each building within the training data set. If you click on a building, you can explore the min, max, median, standard deviation, and the list of units sold.

In general, it seems like the highest standard deviations are in buidings right near the park with very different story height or very few sales in the building. 

```{r, include=FALSE}
building_metrics = train %>% 
    group_by(COOPID, building, lat, lon, ADRNO, ADRSTR) %>% 
    summarize(
        n_sales = n(),
        min = paste0("$", formatC(min(GROSSCON), format = "f", big.mark = ",", digits = 2)), 
        max = paste0("$", formatC(max(GROSSCON), format = "f", big.mark = ",", digits = 2)), 
        median = paste0("$", formatC(quantile(GROSSCON, probs=0.5), format = "f", big.mark = ",", digits = 2)),
        sd = sd(GROSSCON),
        standard_deviation = paste0("$", formatC(sd(GROSSCON), format = "f", big.mark = ",", digits = 2)),
        units = paste(COOPUNITKEY, collapse = ", <br>")
    )
```

```{r}
create_building_map(
    building_metrics,
    display_variables = c("COOPID", "building", "n_sales", "min", "max", "median", "standard_deviation", "units"),
    layers = list(
        list(group="By Standard Deviation", var="sd", color_pal=color_pal_grosscon)
             )
)
```






## Notes {#notes}

The table below summarizes how the tables and maps are shaded.

```{r, echo=FALSE}
# Create the summary data frame
color_pal_summary <- data.frame(
  palette = c(
    "color_pal_r", 
    "color_pal_cod", 
    "color_pal_prd", 
    "color_pal_sales_ratio", 
    "color_pal_grosscon",
    "color_pal_mki"
  ),
  green_range = c(
    "≥ 0.95", 
    "5–10", 
    "0.98–1.03", 
    "0.9–1.1", 
    "N/A",
    "0.9–1.1"
  ),
  below_range = c(
    "0.4–0.95: Red to Yellow to Green", 
    "0–5: Red to Yellow", 
    "0.90–0.98: Dark Blue to Light Blue (Progressive)", 
    "0.6–0.9: Dark Blue to Light Blue (Progressive)", 
    "N/A",
    "0.4–0.9: Red to Yellow"
  ),
  above_range = c(
    "N/A", 
    "10–20: Yellow to Red", 
    "1.03–1.10: Light to Dark Red (Regressive)", 
    "1.1–1.4: Light to Dark Red (Regressive)", 
    "Blue gradient ($200K–$5M)",
    "1.1–1.4: Yellow to Red"
  ),
  stringsAsFactors = FALSE
)

# Display table with minimal formatting
datatable(
  color_pal_summary,
  options = list(
    dom = 't',
    ordering = FALSE
  ),
  rownames = FALSE
)
```
