---
title: "Co-op Sales Model: Comparable Properties"
format: html
execute:
  echo: false
resources:
  - data/Coop/MixedModel/comps_long.zip
---

```{r, include=FALSE}
# read env file to get proxy set up to work
readRenviron("Data/.Renviron")
options(scipen=99999)

library(readxl)
library(stringr)
library(readr)
library(ggplot2)
library(dplyr)
library(sf)
library(rmarkdown)
library(maps)
library(leaflet)
library(scales)
library(DT)
```

```{r functions, include=FALSE, cache=TRUE}
# color palettes to use across plots/maps
color_pal_r <- function(x) {
  sapply(x, function(val) {
    if (is.na(val)) return("#999999")  # gray for NA

    if (val >= 0.95) {
      return("#00FF00")  # pure green, consistent with others
    } else if (val >= 0.7 && val < 0.95) {
      # Green to Yellow: R goes from 0 → 1, G stays at 1
      intensity <- rescale(val, from = c(0.7, 1), to = c(1, 0))  # reverse scale
      return(rgb(intensity, 1, 0))  # green to yellow
    } else if (val >= 0.4 && val < 0.7) {
      # Yellow to Red: G goes from 1 → 0, R stays at 1
      intensity <- rescale(val, from = c(0.4, 0.7), to = c(0, 1))
      return(rgb(1, intensity, 0))  # yellow to red
    } else {
      return("#FF0000")  # solid red for <0.4
    }
  })
}

color_pal_cod <- function(x) {
  sapply(x, function(val) {
    if (is.na(val)) return("#999999")  # gray for NA

    if (val >= 5 & val <= 15) {
      return("#00FF00")  # Still using green for the "ideal" zone
    } else if (val < 5) {
      # Values < 5: interpolate from yellow (at 5) to orange (at 0)
      # Yellow = rgb(1, 1, 0), Orange = rgb(1, 0.5, 0)
      intensity <- scales::rescale(val, from = c(0, 5), to = c(0.5, 1))  # Green component goes from 0.5 to 1
      return(rgb(1, intensity, 0))
    } else {
      # Interpolate from yellow (10) to red (20), capped at red for >20
      capped_val <- min(val, 20)
      intensity <- rescale(capped_val, from = c(10, 20), to = c(1, 0))  # G goes from 1 to 0
      return(rgb(1, intensity, 0))  # Yellow to Red
    }
  })
}

color_pal_prd <- function(x) {
  sapply(x, function(val) {
    if (is.na(val)) return("#999999")  # gray for NA

    if (val >= 0.98 & val <= 1.03) {
      return("#00FF00")  # ideal green
    } else if (val < 0.98) {
      # Light blue (0.98) to dark blue (~0.90): blue stays 1, R & G go from 0.8 to 0
      capped_val <- max(val, 0.90)
      intensity <- scales::rescale(capped_val, from = c(0.90, 0.98), to = c(0, 0.8))
      return(rgb(intensity, intensity, 1))  # light to dark blue
    } else {
      # Light red (1.03) to dark red (~1.10): red stays 1, G & B go from 0.8 to 0
      capped_val <- min(val, 1.10)
      intensity <- scales::rescale(capped_val, from = c(1.03, 1.10), to = c(0.8, 0))
      return(rgb(1, intensity, intensity))  # light to dark red
    }
  })
}

color_pal_sales_ratio <- function(x) {
  sapply(x, function(val) {
    if (is.na(val)) return("#999999")  # gray for NA

    if (val >= 0.9 & val <= 1.1) {
      return("green")  # Ideal range
    } else if (val < 0.9) {
      # Light blue (0.9) to dark blue (0.6): blue stays 1, R & G fade from 0.6 to 0
      capped_val <- max(val, 0.6)
      intensity <- scales::rescale(capped_val, from = c(0.6, 0.9), to = c(0, 0.6))
      return(rgb(intensity, intensity, 1))  # Light blue to dark blue
    } else {
      # Light red (1.1) to dark red (1.4): red stays 1, G & B fade from 0.6 to 0
      capped_val <- min(val, 1.4)
      intensity <- scales::rescale(capped_val, from = c(1.1, 1.4), to = c(0.6, 0))
      return(rgb(1, intensity, intensity))  # Light red to dark red
    }
  })
}


color_pal_grosscon <- function(x){
 
  clamped <- pmin(pmax(x, 200000), 5000000)
    
  pal = colorNumeric(
      palette = "Blues",  # You can also use c("#deebf7", "#3182bd") for more control
      domain = c(200000, 5000000),  # fixed scale
      na.color = "#999999"  # gray for NA
  )
  return(pal(clamped))
}


color_pal_mki <- function(x) {
  sapply(x, function(val) {
    if (is.na(val)) return("#999999")  # gray for NA

    if (val > 1.4) {
      return("#FF0000")  # solid red for extreme high values
    } else if (val > 1.1 && val <= 1.4) {
      # Yellow to Red for values >1.1 to 1.4
      intensity <- rescale(val, from = c(1.1, 1.4), to = c(1, 0))
      return(rgb(1, intensity, 0))  # yellow to red
    } else if (val >= 0.9 && val <= 1.1) {
      return("#00FF00")  # solid yellow as center plateau
    } else if (val >= 0.6 && val < 0.9) {
      # Yellow to Red from 0.9 to 0.6
      intensity <- rescale(val, from = c(0.9, 0.6), to = c(1, 0))
      return(rgb(1, intensity, 0))  # yellow to red
    } else {
      return("#FF0000")  # solid red for <0.6
    }
  })
}

#table formatting
format_metrics_table <- function(data, palettes, opacity = 0.2, exclude_cols = c("ci", "gini")) {
  # Add color columns using provided palettes and opacity
  data <- data %>%
    mutate(across(
      .cols = names(palettes),
      .fns = ~ hex_with_opacity(palettes[[cur_column()]](.x), opacity),
      .names = "color_{.col}"
    ))
  
  # Determine which columns to show
  visible_cols <- setdiff(names(data), c(paste0("color_", names(palettes)), exclude_cols))
  
  # Build the datatable
  dt <- datatable(
    data[, visible_cols],
    options = list(pageLength = 6, scrollX = TRUE)
  )
  
  # Apply conditional formatting
  for (metric in names(palettes)) {
    dt <- dt %>% formatStyle(
      metric,
      backgroundColor = styleEqual(
        data[[metric]],
        data[[paste0("color_", metric)]]
      )
    )
  }
  
  return(dt)
}


# map functions
create_metrics_map <- function(data, by, center_lng = -73.96, center_lat = 40.78, zoom_level = 12) {
  
  leaflet(data = data, height=800) %>%
    addTiles() %>%
    setView(center_lng, center_lat, zoom = zoom_level) %>%
    addProviderTiles("CartoDB.Positron") %>% 
    {
      map <- .
      
          # Create popup content using get() to dynamically access the ID field
          popup_content <- eval(parse(text = paste0(
            "~paste('<strong>", by, ":</strong> ', get('", by, "'), '<br>',",
            "'<strong>n:</strong> ', n, '<br>',",
            "'<strong>median_sales_ratio:</strong> ', median_sales_ratio, '<br>',",
            "'<strong>r:</strong> ', r, '<br>',",
            "'<strong>spearmans:</strong> ', spearmans, '<br>',",
            "'<strong>cod:</strong> ', cod, '<br>',",
            "'<strong>prd:</strong> ', prd, '<br>',",
            "'<strong>prb:</strong> ', prb, '<br>',",
            "'<strong>clapp:</strong> ', clapp, '<br>',",
            "'<strong>ki:</strong> ', ki, '<br>',",
            "'<strong>mki:</strong> ', mki, '<br>')"
          )))
    
      
      # Define function to add polygon layers
      add_polygon_layer <- function(m, group_name, var_name, color_fn) {
        m %>% addPolygons(
          group = group_name,
          fillColor = ~color_fn(get(var_name)),
          fillOpacity = 0.7,
          color = "white",
          weight = 1,
          popup = popup_content
        )
      }
      
      # Add all layers using the function
      map %>%
        add_polygon_layer("By R", "r", color_pal_r) %>%
        add_polygon_layer("By Spearman's Rank", "spearmans", color_pal_r) %>%
        add_polygon_layer("By COD", "cod", color_pal_cod) %>%
        add_polygon_layer("By PRD", "prd", color_pal_prd) %>%
        add_polygon_layer("By MKI", "mki", color_pal_mki)
    } %>%
    
    # Add control to switch between layers
    addLayersControl(
      baseGroups = c("By R", "By Spearman's Rank","By COD", "By PRD", "By MKI"),
      options = layersControlOptions(collapsed = FALSE)
    ) %>%
    
    # Only show one group by default
    hideGroup(c("By Spearman's Rank","By COD", "By PRD","By MKI"))
}

create_point_map <- function(data, 
                             dependent_variables,
                             layers =
                              list(
                                list(group = "By Sale Price", var = "GROSSCON", color_pal = color_pal_grosscon),
                                list(group = "By Predicted Sale Price", var = "prediction", color_pal = color_pal_grosscon),
                                list(group = "By Sales Ratio", var = "sales_ratio", color_pal = color_pal_sales_ratio)
                              ), 
                             center_lng = -73.96, 
                             center_lat = 40.78, 
                             zoom_level = 13) {

  # Required top-level fields, formatted
  popup_content <- sapply(seq_len(nrow(data)), function(i) {
    row <- data[i, ]

    main_info <- paste0(
      "<strong>PID:</strong> ", row[["PID"]], "<br>",
      "<strong>SALEKEY:</strong> ", row[["SALEKEY"]], "<br>",
      "<strong>Sale Price (real value):</strong> $", scales::comma(round(row[["GROSSCON"]])), "<br>",
      "<strong>Predicted Sale Price:</strong> $", scales::comma(round(row[["prediction"]])), "<br>",
      "<strong>Sales Ratio:</strong> ", round(row[["sales_ratio"]], 3), "<br><br>"
    )

    dependent_vars <- paste0(
      "<strong>Dependent Variables:</strong><br>",
      paste0(
        lapply(dependent_variables, function(var) {
          paste0("<strong>", var, ":</strong> ", row[[var]])
        }),
        collapse = "<br>"
      )
    )

    paste0(main_info, dependent_vars)
  })

  map <- leaflet(data = data, height = 800) %>%
    addTiles() %>%
    setView(lng = center_lng, lat = center_lat, zoom = zoom_level) %>%
    addProviderTiles("CartoDB.Positron")

  for (layer in layers) {
    group_name <- layer$group
    var_name <- layer$var
    color_fn <- layer$color_pal

    map <- map %>%
      addCircleMarkers(
        data = data,
        group = group_name,
        fillColor = ~color_fn(get(var_name)),
        color = "black",
        weight = 1,
        radius = 6,
        fillOpacity = 0.7,
        stroke = TRUE,
        popup = popup_content
      )
  }

  map %>%
    addLayersControl(
      baseGroups = sapply(layers, function(l) l$group),
      options = layersControlOptions(collapsed = FALSE)
    ) %>%
    hideGroup(sapply(layers[-1], function(l) l$group))
}


hex_with_opacity <- function(hex, alpha = 1) {
  hex <- gsub("#", "", hex)  # Remove '#' if present

  # Convert alpha (0–1) to 2-digit hex
  alpha_hex <- toupper(sprintf("%02X", round(alpha * 255)))

  paste0("#", hex, alpha_hex)
}

```

```{r data_ingestion, include=FALSE}
source_folder = "Data/Coop/MixedModel"

comps_boro_metrics = readr::read_csv(glue::glue("{source_folder}/comp_metrics.csv")) %>%
    mutate(across(
        .cols = -c(boro, predicted, actual, n, accuracy, uniformity, vertical_equity),
        .fns = ~round(.x, 4)
      )) %>%
    select(-predicted, -actual)

comps_short = read_csv(glue::glue("{source_folder}/comps_short.csv"))

comps_short_trimmed = comps_short %>% 
    mutate(sales_ratio = round(selected_comparable_GROSSCON/GROSSCON,digits = 4)) %>% 
    select(1:4, sales_ratio, 8:10)
```

## Comparable Properties

### Metrics

Now, let's measure the model performance based on the sales price of the property that is most similar to each subject property based on the model's coefficients. Metrics for the model at the city-wide and boro level are displayed below.

***Note:** As of now, both the subject properties and the potential comparable properties are the same training set (80% of all eligible sales)*

```{r, echo=FALSE}
metric_palettes <- list(
  r = color_pal_r,
  spearmans = color_pal_r,
  cod = color_pal_cod,
  prd = color_pal_prd,
  mki = color_pal_mki
)

format_metrics_table(comps_boro_metrics, palettes = metric_palettes, opacity = 0.2)
```

### Comparable Properties (unadjusted)

To examine a specific property's selected comps, search for a COOPUNITKEY to get a short summary in a table format.

```{r, echo=FALSE, warning=FALSE}
datatable(comps_short_trimmed,
          options = list(
            pageLength = 10,     # Show 25 rows per page
            lengthMenu = c(10, 25, 50, 100),
            scrollX = TRUE,      # Horizontal scroll for wide tables
            deferRender = TRUE,  # Speeds up initial load
            dom = 'Blfrtip'      # Add buttons, length menu, filter, etc.
          ),
          filter = "top",        # Adds filter boxes at the top of each column
          rownames = FALSE
) %>% formatStyle(
    # preserves whitespace for copy/paste matching between the table and the map
    columns = c("COOPUNITKEY", "selected_comparable_COOPUNITKEY"),
    `white-space` = "pre"
  )
```

### Interactive Map

Using any COOPUNITKEY listed above, display the comps and their relative features using the interactive map.

```{=html}
<!-- Include Leaflet and D3 -->
<link rel="stylesheet" href="https://unpkg.com/leaflet@1.9.4/dist/leaflet.css" />
<script src="https://unpkg.com/leaflet@1.9.4/dist/leaflet.js"></script>
<script src="https://d3js.org/d3.v7.min.js"></script>
<script src="https://cdnjs.cloudflare.com/ajax/libs/jszip/3.10.1/jszip.min.js"></script>


<!-- UI: Search + Map -->
<div style="margin:1em;"> 
<input type="text" id="unitKeyInput" placeholder="Enter COOPUNITKEY..." style="width: 600px;"/>
<button onclick="searchUnit()">Search</button>
</div>
<div id="map" style="height: 800px;"></div>

<script>
  const map = L.map("map").setView([40.78, -73.96], 13);
  
  const formatDollar = d3.format(",.0f");
  const formatRatio = d3.format(".3f");

  function createPopupContent(d, dependentVars = []) {
  const depDetails = dependentVars.map(v => {
    const value = d[v] ?? "N/A";
    return `<strong>${v}:</strong> ${value}`;
  }).join("<br>");
  
  const similaritySection = (d.similarity_score == null || d.similarity_score === 'NA')
  ? `<strong>Subject Property</strong><br><br>`
  : `<strong>Similiarity to Subject Property:</strong><br>
     <strong>Similarity Score:</strong> ${d.similarity_score}<br>
     <strong>Distance Score:</strong> ${d.distance_score}<br>
     <strong>Variable Score:</strong> ${d.variable_score}<br>
     <strong>Selected as Comparable Property:</strong> ${d.is_selected_comp}<br><br>`;

  return `
    <strong>PID:</strong> ${d.PID}<br>
    <strong>COOPUNITKEY:</strong> ${d.COOPUNITKEY}<br>
    <strong>ADDRESS:</strong> ${d.ADRNO} ${d.ADRSTR}<br>
    <strong>SALEKEY:</strong> ${d.SALEKEY}<br>
    <strong>Sale Price (real value):</strong> $${formatDollar(d.GROSSCON)}<br>
    <strong>Predicted Sale Price:</strong> $${formatDollar(d.prediction)}<br>
    <strong>Sales Ratio:</strong> ${formatRatio(d.sales_ratio)}<br><br>

    ${similaritySection}
    
    <strong>Dependent Variables:</strong><br>${depDetails}
  `;
}

  const dependentVars = ["NEIGHBORHOOD", "BC", "park_dist", "cultural_facility_dist", "subway_dist",
    "affordable","gym", "laundry", "doorman","stories","floor","BEDROOMS","BATHROOMS","log_sqft", "sqft"];

  L.tileLayer("https://{s}.basemaps.cartocdn.com/light_all/{z}/{x}/{y}{r}.png", {
    attribution: '&copy; <a href="https://carto.com/">CARTO</a> | Map data © OpenStreetMap contributors',
    subdomains: "abcd",
    maxZoom: 19
  }).addTo(map);

  let allData = [];
  const markersLayer = L.layerGroup().addTo(map);

  // Load data from a ZIP file instead of CSV
    fetch("data/Coop/MixedModel/comps_long.zip")
      .then(res => res.arrayBuffer())
      .then(JSZip.loadAsync)
      .then(zip => zip.file("comps_long.csv").async("text"))
      .then(text => {
        allData = d3.csvParse(text, d3.autoType).map(d => {
          return d;
        });
      });

  console.log("✅ All data loaded from ZIP:", allData.length);

  // Search function for COOPUNITKEY with partial match
  function searchUnit() {
    const unitKey = document.getElementById("unitKeyInput").value.trim();
    if (!unitKey) return;
    
    console.log("Searching for COOPUNITKEY containing:", unitKey);

    markersLayer.clearLayers();

    // Subject property
    const subject = allData.find(d => 
      d.COOPUNITKEY?.toString().includes(unitKey) && d.is_subject_property === "TRUE"
    );
    
    if (!subject) {
      alert("Subject property not found.");
      return;
    }

    // Add subject marker (red)
    L.circleMarker([+subject.lat, +subject.lon], {
      radius: 8,
      color: "black",
      fillColor: "blue",
      fillOpacity: 0.9
    }).addTo(markersLayer)
    .bindPopup(createPopupContent(subject, dependentVars))
    .openPopup()

    map.setView([+subject.lat, +subject.lon], 18);

    // Add comps (same COOPUNITKEY, is_subject_property === false)
    const comps = allData.filter(d => 
      d.subject_COOPUNITKEY?.toString().includes(unitKey) && d.is_subject_property === "FALSE"
    );
    
    console.log("✅ Comps loaded:", comps.length); // <-- Check number of rows loaded

    comps.forEach((d, i) => {
      const offset = 0.0001; // ~10 meters
      const angle = (i / comps.length) * 2 * Math.PI;
      const jitterLat = +d.lat + offset * Math.sin(angle);
      const jitterLon = +d.lon + offset * Math.cos(angle);
    
      const isSelected = d.is_selected_comp === "TRUE";
      
      L.circleMarker([jitterLat, jitterLon], {
        radius:  d.is_selected_comp === "TRUE" ? 8 : 6,
        color: d.is_selected_comp === "TRUE" ? "black" : "gray",
        fillColor: d.is_selected_comp === "TRUE" ? "green" : "gray",
        fillOpacity: 0.9
      }).addTo(markersLayer)
      .bindPopup(createPopupContent(d, dependentVars));
    });
  }
</script>
```
