---
title: "Condo Sales Model"

repo-url: https://github.com/nycdepartmentoffinance/nycdepartmentoffinance.github.io
---

```{r, include=FALSE}

library(readxl)
library(stringr)
library(readr)
library(ggplot2)
library(dplyr)
library(sf)
library(maps)
library(leaflet)
library(scales)
library(DT)
# read env file to get proxy set up to work
readRenviron("Data/.Renviron")
options(scipen=99999)

citywidedecs <- read_xlsx("data/Condo/CitywideMetricsHalves.xlsx",sheet = "Citywide", range = cell_rows(1:12)) %>%
  mutate(across(
        .cols = -c(grosscondecile, cutoff, predicted, actual, n, accuracy, uniformity, vertical_equity), 
        .fns = ~round(.x, 4)
      ))
half1decs <- read_xlsx("data/Condo/CitywideMetricsHalves.xlsx",sheet = "Half1", range = cell_rows(1:12)) %>%
    mutate(across(
        .cols = -c(grosscondecile, cutoff, predicted, actual, n, accuracy, uniformity, vertical_equity), 
        .fns = ~round(.x, 4)
      )) %>% 
    select(-predicted, -actual) %>%
  relocate(cutoff,.after = grosscondecile)
half2decs <- read_xlsx("data/Condo/CitywideMetricsHalves.xlsx",sheet = "Half2", range = cell_rows(1:12)) %>%
    mutate(across(
        .cols = -c(grosscondecile, cutoff, predicted, actual, n, accuracy, uniformity, vertical_equity), 
        .fns = ~round(.x, 4)
      )) %>% 
    select(-predicted, -actual) %>%
  relocate(cutoff,.after = grosscondecile)
citywideborodecs <- read_xlsx("data/Condo/CitywideMetricsHalves.xlsx",sheet = "Citywide", range = cell_rows(14:65)) %>%
    mutate(across(
        .cols = -c(boro, decile, predicted, actual, n, accuracy, uniformity, vertical_equity,clapp), 
        .fns = ~round(.x, 4)
      )) %>% 
    select(-predicted, -actual) 
half1borodecs <- read_xlsx("data/Condo/CitywideMetricsHalves.xlsx",sheet = "Half1", range = cell_rows(14:65)) %>%
    mutate(across(
        .cols = -c(boro, decile, predicted, actual, n, accuracy, uniformity, vertical_equity,clapp), 
        .fns = ~round(.x, 4)
      )) %>% 
    select(-predicted, -actual)
half2borodecs <- read_xlsx("data/Condo/CitywideMetricsHalves.xlsx",sheet = "Half2", range = cell_rows(14:65)) %>%
  mutate(r = as.numeric(r)) %>%
    mutate(across(
        .cols = -c(boro, decile, predicted, actual, n, accuracy, uniformity, vertical_equity,clapp), 
        .fns = ~round(.x, 4)
      )) %>% 
    select(-predicted, -actual) 
citywidenbhddecs <- read_xlsx("data/Condo/CitywideMetricsHalves.xlsx",sheet = "Citywide", na = "NA", range = cell_rows(67:1083))%>%
    mutate(r = as.numeric(r)) %>%
    mutate(across(
        .cols = -c(boro, decile, nbhd, cutoff, median_sales_ratio, predicted, actual, n, accuracy, uniformity, vertical_equity), 
        .fns = ~round(.x, 4)
      )) %>% 
    select(-predicted, -actual) %>%
  relocate(cutoff,.after = decile)
half1nbhddecs <- read_xlsx("data/Condo/CitywideMetricsHalves.xlsx",sheet = "Half1",na = "NA", range = cell_rows(67:958)) %>%
    mutate(across(
        .cols = -c(boro, decile,  nbhd,cutoff, predicted, actual,median_sales_ratio, n, accuracy, uniformity, vertical_equity), 
        .fns = ~round(.x, 4)
      )) %>% 
    select(-predicted, -actual) %>%
  relocate(cutoff,.after = decile)
half2nbhddecs <- read_xlsx("data/Condo/CitywideMetricsHalves.xlsx",sheet = "Half2",na = "NA", range = cell_rows(67:1278)) %>%
    mutate(across(
        .cols = -c(boro, decile,  nbhd,cutoff, predicted, actual,median_sales_ratio, n, accuracy, uniformity, vertical_equity), 
        .fns = ~round(.x, 4)
      )) %>% 
    select(-predicted, -actual) %>%
  relocate(cutoff,.after = decile)

basedata <- read_csv("data/Condo/UAFElevBCATJoin.csv")
nbhdmetrics <- read.csv("Data/Condo/NBHDMetrics.csv")

decilemetrics <- read.csv("Data/Condo/DecileMetrics.csv") %>% 
    mutate(across(
        .cols = -c(boro, grosscondecile, cutoff, predicted, actual, n, accuracy, uniformity, vertical_equity), 
        .fns = ~round(.x, 4)
      )) %>% 
    select(-predicted, -actual) %>%
  relocate(cutoff,.after = grosscondecile)

```

```{r functions, include=FALSE, cache=TRUE}
# color palettes to use across plots/maps
color_pal_r <- function(x) {
  sapply(x, function(val) {
    if (is.na(val)) return("#999999")  # gray for NA

    if (val >= 0.95) {
      return("#00FF00")  # pure green, consistent with others
    } else if (val >= 0.7 && val < 0.95) {
      # Green to Yellow: R goes from 0 → 1, G stays at 1
      intensity <- rescale(val, from = c(0.7, 1), to = c(1, 0))  # reverse scale
      return(rgb(intensity, 1, 0))  # green to yellow
    } else if (val >= 0.4 && val < 0.7) {
      # Yellow to Red: G goes from 1 → 0, R stays at 1
      intensity <- rescale(val, from = c(0.4, 0.7), to = c(0, 1))
      return(rgb(1, intensity, 0))  # yellow to red
    } else {
      return("#FF0000")  # solid red for <0.4
    }
  })
}

color_pal_cod <- function(x) {
  sapply(x, function(val) {
    if (is.na(val)) return("#999999")  # gray for NA

    if (val >= 5 & val <= 10) {
      return("#00FF00")  # Still using green for the "ideal" zone
    } else if (val < 5) {
      # Values < 5: interpolate from yellow (at 5) to orange (at 0)
      # Yellow = rgb(1, 1, 0), Orange = rgb(1, 0.5, 0)
      intensity <- scales::rescale(val, from = c(0, 5), to = c(0.5, 1))  # Green component goes from 0.5 to 1
      return(rgb(1, intensity, 0))
    } else {
      # Interpolate from yellow (10) to red (20), capped at red for >20
      capped_val <- min(val, 20)
      intensity <- rescale(capped_val, from = c(10, 20), to = c(1, 0))  # G goes from 1 to 0
      return(rgb(1, intensity, 0))  # Yellow to Red
    }
  })
}

color_pal_prd <- function(x) {
  sapply(x, function(val) {
    if (is.na(val)) return("#999999")  # gray for NA

    if (val >= 0.98 & val <= 1.03) {
      return("#00FF00")  # ideal green
    } else if (val < 0.98) {
      # Light blue (0.98) to dark blue (~0.90): blue stays 1, R & G go from 0.8 to 0
      capped_val <- max(val, 0.90)
      intensity <- scales::rescale(capped_val, from = c(0.90, 0.98), to = c(0, 0.8))
      return(rgb(intensity, intensity, 1))  # light to dark blue
    } else {
      # Light red (1.03) to dark red (~1.10): red stays 1, G & B go from 0.8 to 0
      capped_val <- min(val, 1.10)
      intensity <- scales::rescale(capped_val, from = c(1.03, 1.10), to = c(0.8, 0))
      return(rgb(1, intensity, intensity))  # light to dark red
    }
  })
}

color_pal_sales_ratio <- function(x) {
  sapply(x, function(val) {
    if (is.na(val)) return("#999999")  # gray for NA

    if (val >= 0.9 & val <= 1.1) {
      return("green")  # Ideal range
    } else if (val < 0.9) {
      # Light blue (0.9) to dark blue (0.6): blue stays 1, R & G fade from 0.6 to 0
      capped_val <- max(val, 0.6)
      intensity <- scales::rescale(capped_val, from = c(0.6, 0.9), to = c(0, 0.6))
      return(rgb(intensity, intensity, 1))  # Light blue to dark blue
    } else {
      # Light red (1.1) to dark red (1.4): red stays 1, G & B fade from 0.6 to 0
      capped_val <- min(val, 1.4)
      intensity <- scales::rescale(capped_val, from = c(1.1, 1.4), to = c(0.6, 0))
      return(rgb(1, intensity, intensity))  # Light red to dark red
    }
  })
}


color_pal_grosscon <- function(x){
 
  clamped <- pmin(pmax(x, 200000), 5000000)
    
  pal = colorNumeric(
      palette = "Blues",  # You can also use c("#deebf7", "#3182bd") for more control
      domain = c(200000, 5000000),  # fixed scale
      na.color = "#999999"  # gray for NA
  )
  return(pal(clamped))
}

#table formatting

format_metrics_table <- function(data, palettes, opacity = 0.2, exclude_cols = c("ci", "gini")) {
  # Add color columns using provided palettes and opacity
  data <- data %>%
    mutate(across(
      .cols = names(palettes),
      .fns = ~ hex_with_opacity(palettes[[cur_column()]](.x), opacity),
      .names = "color_{.col}"
    ))
  
  # Determine which columns to show
  visible_cols <- setdiff(names(data), c(paste0("color_", names(palettes)), exclude_cols))
  
  # Build the datatable
  dt <- datatable(
    data[, visible_cols],
    options = list(pageLength = 6, scrollX = TRUE)
  )
  
  # Apply conditional formatting
  for (metric in names(palettes)) {
    dt <- dt %>% formatStyle(
      metric,
      backgroundColor = styleEqual(
        data[[metric]],
        data[[paste0("color_", metric)]]
      )
    )
  }
  
  return(dt)
}


format_decile_metrics_table <- function(data, palettes, opacity = 0.2, exclude_cols = c("ci", "gini")) {
  # Add color columns using provided palettes and opacity
  data <- data %>%
    mutate(across(
      .cols = names(palettes),
      .fns = ~ hex_with_opacity(palettes[[cur_column()]](.x), opacity),
      .names = "color_{.col}"
    ))
  
  # Determine which columns to show
  visible_cols <- setdiff(names(data), c(paste0("color_", names(palettes)), exclude_cols))
  
  # Build the datatable
  dt <- datatable(
    data[, visible_cols],
    options = list(pageLength = 11, scrollX = TRUE)
  )
  
  # Apply conditional formatting
  for (metric in names(palettes)) {
    dt <- dt %>% formatStyle(
      metric,
      backgroundColor = styleEqual(
        data[[metric]],
        data[[paste0("color_", metric)]]
      )
    )
  }
  
  return(dt)
}

# map functions
create_metrics_map <- function(data, by, center_lng = -73.96, center_lat = 40.78, zoom_level = 12) {
  
  leaflet(data = data, height=800) %>%
    addTiles() %>%
    setView(center_lng, center_lat, zoom = zoom_level) %>%
    addProviderTiles("CartoDB.Positron") %>% 
    {
      map <- .
      
          # Create popup content using get() to dynamically access the ID field
          popup_content <- eval(parse(text = paste0(
            "~paste('<strong>", by, ":</strong> ', get('", by, "'), '<br>',",
            "'<strong>n:</strong> ', n, '<br>',",
            "'<strong>median_sales_ratio:</strong> ', median_sales_ratio, '<br>',",
            "'<strong>r:</strong> ', r, '<br>',",
            "'<strong>spearmans:</strong> ', spearmans, '<br>',",
            "'<strong>cod:</strong> ', cod, '<br>',",
            "'<strong>prd:</strong> ', prd, '<br>',",
            "'<strong>prb:</strong> ', prb, '<br>',",
            "'<strong>clapp:</strong> ', clapp, '<br>',",
            "'<strong>ki:</strong> ', ki, '<br>',",
            "'<strong>mki:</strong> ', mki, '<br>')"
          )))
    
      
      # Define function to add polygon layers
      add_polygon_layer <- function(m, group_name, var_name, color_fn) {
        m %>% addPolygons(
          group = group_name,
          fillColor = ~color_fn(get(var_name)),
          fillOpacity = 0.7,
          color = "white",
          weight = 1,
          popup = popup_content
        )
      }
      
      # Add all layers using the function
      map %>%
        add_polygon_layer("By R", "r", color_pal_r) %>%
        add_polygon_layer("By Spearman's Rank", "spearmans", color_pal_r) %>%
        add_polygon_layer("By COD", "cod", color_pal_cod) %>%
        add_polygon_layer("By PRD", "prd", color_pal_prd) %>%
        add_polygon_layer("By MKI", "mki", color_pal_r)
    } %>%
    
    # Add control to switch between layers
    addLayersControl(
      baseGroups = c("By R", "By Spearman's Rank","By COD", "By PRD", "By MKI"),
      options = layersControlOptions(collapsed = FALSE)
    ) %>%
    
    # Only show one group by default
    hideGroup(c("By Spearman's Rank","By COD", "By PRD","By MKI"))
}

create_point_map <- function(data, 
                             dependent_variables,
                             layers =
                              list(
                                list(group = "By Sale Price", var = "GROSSCON", color_pal = color_pal_grosscon),
                                list(group = "By Predicted Sale Price", var = "prediction", color_pal = color_pal_grosscon),
                                list(group = "By Sales Ratio", var = "sales_ratio", color_pal = color_pal_sales_ratio)
                              ), 
                             center_lng = -73.96, 
                             center_lat = 40.78, 
                             zoom_level = 13) {

  # Required top-level fields, formatted
  popup_content <- sapply(seq_len(nrow(data)), function(i) {
    row <- data[i, ]

    main_info <- paste0(
      "<strong>PID:</strong> ", row[["PID"]], "<br>",
      "<strong>SALEKEY:</strong> ", row[["SALEKEY"]], "<br>",
      "<strong>Sale Price (real value):</strong> $", scales::comma(round(row[["GROSSCON"]])), "<br>",
      "<strong>Predicted Sale Price:</strong> $", scales::comma(round(row[["prediction"]])), "<br>",
      "<strong>Sales Ratio:</strong> ", round(row[["sales_ratio"]], 3), "<br><br>"
    )

    dependent_vars <- paste0(
      "<strong>Dependent Variables:</strong><br>",
      paste0(
        lapply(dependent_variables, function(var) {
          paste0("<strong>", var, ":</strong> ", row[[var]])
        }),
        collapse = "<br>"
      )
    )

    paste0(main_info, dependent_vars)
  })

  map <- leaflet(data = data, height = 800) %>%
    addTiles() %>%
    setView(lng = center_lng, lat = center_lat, zoom = zoom_level) %>%
    addProviderTiles("CartoDB.Positron")

  for (layer in layers) {
    group_name <- layer$group
    var_name <- layer$var
    color_fn <- layer$color_pal

    map <- map %>%
      addCircleMarkers(
        data = data,
        group = group_name,
        fillColor = ~color_fn(get(var_name)),
        color = "black",
        weight = 1,
        radius = 6,
        fillOpacity = 0.7,
        stroke = TRUE,
        popup = popup_content
      )
  }

  map %>%
    addLayersControl(
      baseGroups = sapply(layers, function(l) l$group),
      options = layersControlOptions(collapsed = FALSE)
    ) %>%
    hideGroup(sapply(layers[-1], function(l) l$group))
}


hex_with_opacity <- function(hex, alpha = 1) {
  hex <- gsub("#", "", hex)  # Remove '#' if present

  # Convert alpha (0–1) to 2-digit hex
  alpha_hex <- toupper(sprintf("%02X", round(alpha * 255)))

  paste0("#", hex, alpha_hex)
}

```

```{r data_ingestion, include=FALSE}

source_folder = "Data/Condo"

# 
# # model metrics
# boro_metrics = readr::read_csv(glue::glue("{source_folder}/borometrics.csv")) %>% 
#     mutate(across(
#         .cols = -c(boro, predicted, actual, n, accuracy, uniformity, vertical_equity), 
#         .fns = ~round(.x, 4)
#       )) %>% 
#     select(-predicted, -actual)

neighborhood_metrics = readr::read_csv(glue::glue("{source_folder}/NBHDNameJoin.csv")) 
# 
# block_metrics = readr::read_csv(glue::glue("{source_folder}/manhattan_block_metrics.csv")) %>% 
#     mutate(block = as.numeric(block))

#train_df = readr::read_csv(glue::glue("{source_folder}/train_with_yhat.csv"))

# shapefiles // geographic transformations

neighborhood_shapefile = sf::st_read("Data/DOF_NBHD_Dissolved/DOF_NBHD_Dissolved.shp") %>% 
    st_transform(crs=4326) %>%
    left_join(neighborhood_metrics, by=c("CODE" = "nbhd")) %>%

  filter(!is.na(n)) %>%
  filter(n!=1)

#block_shapefile = sf::st_read("Data/TAX_BLOCK_POLYGON_LAYER/TAX_BLOCK_POLYGON_LAYER.shp") %>% 
    # st_transform(crs=4326) %>% 
    # filter(BORO == "1") %>% 
    # select(BLOCK, geometry) %>% 
    # left_join(block_metrics, by=c("BLOCK"="block"))

#train <- sf::st_as_sf(train_df, coords= c("xcoord","ycoord"), crs=2263) %>% 
  #  st_transform(crs=4326) 
```

# Mixed Model

Using sales from 2022-2024 in Manhattan, Brooklyn, and Queens, and sales from 2020-2024 in the Bronx and Staten Island to predict the sale price of condos across NYC, the most successful mixed model (or hierarchical linear model) to date is the following:

```{r, results='hide'}
logtagrosscon ~ nbhd  + X + Y  + trend + UAF + elev +
            (1 + STY  | BLOCK) + (1 + logsf + floor_extract| ADDRESS)

```

## Citywide Model Deciles

These tables show the metrics divided by decile on a model run on the entire city population simultaenously, rather than boro-by-boro. Below are deciles overall, deciles by boro, and deciles by nbhd. 

### Citywide Overall Deciles

```{r, echo=FALSE}
metric_palettes <- list(
  r = color_pal_r,
  spearmans = color_pal_r,
  cod = color_pal_cod,
  prd = color_pal_prd,
  mki = color_pal_r
)

  
format_decile_metrics_table(citywidedecs, palettes = metric_palettes, opacity = 0.2)

```
### Citywide Model Deciles by Boro

#### Manhattan
```{r, echo=FALSE}
mhmetrics <- citywideborodecs %>%
  filter(boro == 1) %>%
  select(-boro)
  
format_decile_metrics_table(mhmetrics, palettes = metric_palettes, opacity = 0.2)
```

#### The Bronx
```{r, echo=FALSE}
bxmetrics <- citywideborodecs %>%
  filter(boro == 2) %>%
  select(-boro)
  
format_decile_metrics_table(bxmetrics, palettes = metric_palettes, opacity = 0.2)
```

#### Brooklyn
```{r, echo=FALSE}
bkmetrics <- citywideborodecs %>%
  filter(boro == 3) %>%
  select(-boro)
  
format_decile_metrics_table(bkmetrics, palettes = metric_palettes, opacity = 0.2)
```

#### Queens
```{r, echo=FALSE}
qnmetrics <- citywideborodecs %>%
  filter(boro == 4) %>%
  select(-boro)
  
format_decile_metrics_table(qnmetrics, palettes = metric_palettes, opacity = 0.2)
```

#### Staten Island

```{r, echo=FALSE}
simetrics <- citywideborodecs %>%
  filter(boro == 5) %>%
  select(-boro)
  
format_decile_metrics_table(simetrics, palettes = metric_palettes, opacity = 0.2)
```

### Citywide Model Deciles By NBHD

The following shows the metrics broken down by decile for each nbhd in the city. 

```{r, echo=FALSE}
format_decile_metrics_table(citywidenbhddecs, palettes = metric_palettes, opacity = 0.2)
```
### Citywide Split

In order to determine whether the poorer metrics were a result of sampling error or poor model calibration, the population was randomly split in half, and the models and metrics were re-run for each half. 



### Split 1 Overall Deciles

```{r, echo=FALSE}
metric_palettes <- list(
    r = color_pal_r,
    spearmans = color_pal_r,
    cod = color_pal_cod,
    prd = color_pal_prd,
    mki = color_pal_r
)


format_decile_metrics_table(half1decs, palettes = metric_palettes, opacity = 0.2)

```
### Split 1 Model Deciles by Boro

#### Manhattan
```{r, echo=FALSE}
mhmetrics <- half1borodecs %>%
    filter(boro == 1) %>%
    select(-boro)

format_decile_metrics_table(mhmetrics, palettes = metric_palettes, opacity = 0.2)
```

#### The Bronx
```{r, echo=FALSE}
bxmetrics <- half1borodecs %>%
    filter(boro == 2) %>%
    select(-boro)

format_decile_metrics_table(bxmetrics, palettes = metric_palettes, opacity = 0.2)
```

#### Brooklyn
```{r, echo=FALSE}
bkmetrics <- half1borodecs %>%
    filter(boro == 3) %>%
    select(-boro)

format_decile_metrics_table(bkmetrics, palettes = metric_palettes, opacity = 0.2)
```

#### Queens
```{r, echo=FALSE}
qnmetrics <- half1borodecs %>%
    filter(boro == 4) %>%
    select(-boro)

format_decile_metrics_table(qnmetrics, palettes = metric_palettes, opacity = 0.2)
```

#### Staten Island

```{r, echo=FALSE}
simetrics <- half1borodecs %>%
    filter(boro == 5) %>%
    select(-boro)

format_decile_metrics_table(simetrics, palettes = metric_palettes, opacity = 0.2)
```

### Split 1 Model Deciles By NBHD


```{r, echo=FALSE}
format_decile_metrics_table(half1nbhddecs, palettes = metric_palettes, opacity = 0.2)
```


### Split 2 Overall Deciles

```{r, echo=FALSE}
metric_palettes <- list(
    r = color_pal_r,
    spearmans = color_pal_r,
    cod = color_pal_cod,
    prd = color_pal_prd,
    mki = color_pal_r
)


format_decile_metrics_table(half2decs, palettes = metric_palettes, opacity = 0.2)

```
### Split 2 Model Deciles by Boro

#### Manhattan
```{r, echo=FALSE}
mhmetrics <- half2borodecs %>%
    filter(boro == 1) %>%
    select(-boro)

format_decile_metrics_table(mhmetrics, palettes = metric_palettes, opacity = 0.2)
```

#### The Bronx
```{r, echo=FALSE}
bxmetrics <- half2borodecs %>%
    filter(boro == 2) %>%
    select(-boro)

format_decile_metrics_table(bxmetrics, palettes = metric_palettes, opacity = 0.2)
```

#### Brooklyn
```{r, echo=FALSE}
bkmetrics <- half2borodecs %>%
    filter(boro == 3) %>%
    select(-boro)

format_decile_metrics_table(bkmetrics, palettes = metric_palettes, opacity = 0.2)
```

#### Queens
```{r, echo=FALSE}
qnmetrics <- half2borodecs %>%
    filter(boro == 4) %>%
    select(-boro)

format_decile_metrics_table(qnmetrics, palettes = metric_palettes, opacity = 0.2)
```

#### Staten Island

```{r, echo=FALSE}
simetrics <- half2borodecs %>%
    filter(boro == 5) %>%
    select(-boro)

format_decile_metrics_table(simetrics, palettes = metric_palettes, opacity = 0.2)
```

### Split 2 Model Deciles By NBHD

```{r, echo=FALSE}
format_decile_metrics_table(half2nbhddecs, palettes = metric_palettes, opacity = 0.2)
```

## Boro-Specific

The interactive map below visualizes key assessment accuracy metrics at the neighborhood level across New York City. Toggle between the different layers which shade neighborhoods by different performance indicators, such as: R, Spearman's Rank, Coefficient of Dispersion (COD), Price-related Differential (PRD), and Modified Kakwani Index (MKI). Clicking on a neighborhood reveals a popup with detailed statistics for that area.

This map displays every neighborhood with at least two qualified sales in the boro-specific time period. While the same model was applied to every borough, it was done so individually, and not as a citywide population.

```{r}

create_metrics_map(
  data = neighborhood_shapefile, 

  by = "FullNbhdName",

  center_lng = -73.99, 
  center_lat = 40.76, 
  zoom_level = 11
)
```

<br>
This map only displays neighborhoods with at least 10 qualified sales in the boro-specific period.

```{r,include=FALSE}
min10 <- neighborhood_shapefile %>%
  filter(n>=10)
```


```{r}
create_metrics_map(
  data = min10, 
  by = "FullNbhdName",

  center_lng = -73.99, 
  center_lat = 40.76, 
  zoom_level = 11
)
```

## Metrics by Sale Price Decile/Quintile
<br> 
The following displays the metrics  for each borough divided by sale price quintile (in Staten Island) or decile (for all other boroughs).


### Manhattan

```{r, echo=FALSE}
metric_palettes <- list(
  r = color_pal_r,
  spearmans = color_pal_r,
  cod = color_pal_cod,
  prd = color_pal_prd,
  mki = color_pal_r
)

mhmetrics <- decilemetrics %>%
  filter(boro == 1) %>%
  select(-boro)
  
format_decile_metrics_table(mhmetrics, palettes = metric_palettes, opacity = 0.2)

```

### The Bronx

```{r, echo=FALSE}

bxmetrics <- decilemetrics %>%
  filter(boro == 2) %>%
  select(-boro)
  
format_decile_metrics_table(bxmetrics, palettes = metric_palettes, opacity = 0.2)

```

### Brooklyn

```{r, echo=FALSE}

bkmetrics <- decilemetrics %>%
  filter(boro == 3) %>%
  select(-boro)
  
format_decile_metrics_table(bkmetrics, palettes = metric_palettes, opacity = 0.2)
```



### Queens

```{r, echo=FALSE}

qnmetrics <- decilemetrics %>%
  filter(boro == 4) %>%
  select(-boro)
  
format_decile_metrics_table(qnmetrics, palettes = metric_palettes, opacity = 0.2)
```
### Staten Island

```{r, echo=FALSE}


simetrics <- decilemetrics %>%
  filter(boro == 5) %>%
  select(-boro)
  
format_metrics_table(simetrics, palettes = metric_palettes, opacity = 0.2)
```

## Notes


The table below summarizes how the tables and maps are shaded.

```{r, echo=FALSE}
# Create the summary data frame
color_pal_summary <- data.frame(
  palette = c(
    "color_pal_r", 
    "color_pal_cod", 
    "color_pal_prd", 
    "color_pal_sales_ratio", 
    "color_pal_grosscon"
  ),
  green_range = c(
    "≥ 0.95", 
    "5–10", 
    "0.98–1.03", 
    "0.9–1.1", 
    "N/A"
  ),
  below_range = c(
    "0.4–0.95: Red to Yellow to Green", 
    "0–5: Red to Yellow", 
    "0.90–0.98: Dark Blue to Light Blue (Progressive)", 
    "0.6–0.9: Dark Blue to Light Blue (Progressive)", 
    "N/A"
  ),
  above_range = c(
    "N/A", 
    "10–20: Yellow to Red", 
    "1.03–1.10: Light to Dark Red (Regressive)", 
    "1.1–1.4: Light to Dark Red (Regressive)", 
    "Blue gradient ($200K–$5M)"
  ),
  stringsAsFactors = FALSE
)

# Display table with minimal formatting
datatable(
  color_pal_summary,
  options = list(
    dom = 't',
    ordering = FALSE
  ),
  rownames = FALSE
)

```
